---
title: "Multivariate Regression"
format: 
  revealjs:
    slide-number: c/t
editor: visual
---

## Multivariate Regression

-   in the last lecture we considered the problem of simple linear regression, where we had one $x$ variable and a response, $y$, that we want to try to understand in the context of $x$
-   now we want to consider a much more complex situation, one where there are many (sometimes millions or tens of millions)

## Multiple linear regression

-   a per case model $$
    y_i = \beta_0 + \beta_1 \cdot x_1 + \ldots + \beta_k \cdot x_k + \epsilon_i
    $$
-   a matrix/vector version of the model
-   in that model we have $n$ observations on $k$ variables, so $\bf{X}$ is an $n \times k$ matrix

$$
  \bf{Y} = \bf{X}^\prime \bf{\beta} + \bf{\epsilon}
$$

## Data

-   for this section we will consider a portion of a very large complex data set
-   the National Health and Nutrition Examination Survey (NHANES)
-   designed to assess the health and nutritional status of adults and children in the United States
-   done in two year cycles since 1999
-   we will look at a small proportion of the data from the 2015-2016 cycle

## Some R code

Loading up some data from

```{r}
load("d4.rda")
lm1 = lm(d4$LBXTC ~ d4$RIDAGEYR)
summary(lm1)

plot(lm1$fitted.values, lm1$residuals)
lines(x=sort(lm1$fitted.values), y=pl, col="blue", lwd=2)
abline(h=0, col="red")



```

## Description of data

For Variables: <https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BPQ_I.htm>

<https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DIQ_I.htm>

```{r}
#| eval: false

#  Variables and the tables they come from:
# this code needs to run in the NHANES docker container then the data are exported to here

  library("phonto")
  cols =   list( DEMO_I=c("RIDAGEYR","RIAGENDR","RIDRETH1","DMDEDUC2","years",
                           "WTINT2YR", "SDMVPSU", "SDMVSTRA", "INDFMPIR"),
            BPQ_I=c('BPQ050A','BPQ020','BPQ080','BPQ100D'),
            HDL_I=c("LBDHDD"),
            GHB_I="LBXGH",
            DIQ_I=c("DIQ010","DIQ050","DIQ070","DIQ160"),
            BMX_I="BMXBMI",
            TCHOL_I="LBXTC")
 data = jointQuery(cols)

 d2 = nhanesTranslate("DEMO_I", colnames=c("RIAGENDR","RIDRETH1","DMDEDUC2"), data =data)
 d3 = nhanesTranslate("BPQ_I", colnames=c('BPQ050A','BPQ020','BPQ080','BPQ100D'), data=d2)
 d4 = nhanesTranslate("DIQ_I", colnames=c("DIQ010","DIQ050","DIQ070","DIQ160"), data=d3)

 save(d4, file="/mnt/projects/d4.rda", compress="xz")


```
