---
title: "BDS^3 2023 applicants selection"
author: "Wolfgang Huber"
format: 
  html:
   df-print: paged
   embed-resources: true
---


```{r}
#| label: global_options
#| echo: false
knitr::opts_chunk$set(cache = FALSE)
options(error = recover)
```

```{r}
#| label: setup
#| message: FALSE
library("dplyr")
library("ggplot2")
```

```{r}
#| label: read
#| message: FALSE
x = readr::read_csv("BDS^3 Applications - Form responses 1.csv")
x = dplyr::filter(x, !grepl("Dnipro", x$`Fedya score`) & !is.na(Timestamp))
dim(x)
colnames(x) = sub(" score", "", colnames(x))
x = mutate(x, 
  Helena   = as.numeric(sub("1/2", "1.5", sub("0/1", "0.5", Helena))),
  Julia    = as.numeric(Julia),
  Wolfgang = as.numeric(Wolfgang)
)
with(x, table(Julia, Helena, useNA = "ifany"))
otherscores = c("Wolfgang", "Roderic", "Fedya", "Lada")
lapply(otherscores, function(j) with(x, table(Julia, Helena, get(j)))) |> setNames(otherscores)
```

Some scoring arithmetics. The first row (value 100) are the ones definitively selected.

```{r}
x$select = rep(0, nrow(x))
x$select[(x$Julia>=1 & x$Helena>=0)   | (x$Julia>=0 & x$Helena>=0.5)] =  30
x$select[(x$Julia>=1 & x$Helena>=0.5) | (x$Julia>=0 & x$Helena>=1)  ] = 100 
x$select[ x$select == 30 & x$Wolfgang > 0 ] = 60

table(x$select, useNA = "ifany")
```
We can invite the `r sum(x$select==100, na.rm = TRUE)` with score 100 based on Helena's and Julia's scores right away.
For those that got 30 based on Helena's and Julia's scores (Julia 1, Helena 0; Julia 0, Helena 0.5), Wolfgang had a look, and suggests picking another `r sum(x$select==60, na.rm = TRUE)`.

Let's make sure we give a nice explanation  to those we cannot invite. ("lots of excellent applications, unfortunately we had limited number of places, hoping for you it works another time, please don't give up.")

```{r}
#| results: markup 
fn = sprintf("Selection %s", format(Sys.time()))
openxlsx2::write_xlsx(
  filter(x, select >= 60) |> arrange(-select), 
  file = paste0(fn, ".xlsx"), colNames = TRUE, sheetName = fn, creator = "Wolfgang Huber"
)
```

```{r}
filter(x, is.na(select))
```
