---
title: "Hypothesis_testing"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")

```

1. Clone, or update your existing clone of, the GitHub repository https://github.com/Huber-group-EMBL/BDS3 

2. read the single cell count data.

```{r}
perez = read.csv("data/Perez_2022_cell_type_counts.csv")
head(perez)
```

3. Explore the following columns in the data: `SampleID`, `Age`, `Age_group` and `manual_anno2`. How many different unique values do they have? 

```{r}
table(perez$SampleID) # what do you think this function is doing?
```

4. We want to quantify change in the abundance of the following populations with age: `MF-I`, `MF-II`, `MYH8+`, `MuSC`. Subset the data to these populations. Start by calculating the number of cells of each type in every sample.  

```{r}

```


4. Do a t-test to test for changes in the abundance with age using `Age` as a binary covariate. What do you observe? Which populations are significantly differentially abundant with age?

```{r}

```

5. Fit a linear model to quantify change in the abundance with age using `Age` as a continous covariate. How do the results compare?

```{r}

```


