---
title: "Hypothesis_testing"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")

```

1. git clone repository https://github.com/Huber-group-EMBL/BDS3 again 

2. read into the cell type count data.
```{r}
#setwd("/home/jovyan/BDS3/")
data_path = "BDS3/lab/03-singlecell/data/Perez_2022_cell_type_counts.csv"
data = read.csv("/home/jovyan/BDS3/lab/03-singlecell/data/Perez_2022_cell_type_counts.csv")

```

3. Explore the following columns in the data: 'SampleID', 'Age', 'Age_group' and 'manual_anno2'. How many different unique values they have? 
```{r}
table(data$SampleID) # what do you think this function is doing?

```

4. We want to quantify change in the abundance of the following populations with age: "MF-I", "MF-II", "MYH8+", "MuSC".Subset the data to the following populations. Start by caculating number of cells of each type in every sample.  
```{r}

```


4. Run the linear model to quantify change in the abundance with age using Age as a binary covariate. What do you observe? Which populations are significantly differentially abudnant with age? 
```{r}

```


